\documentclass[11pt,letter]{article}
\usepackage[top=1.00in, bottom=1.0in, left=1.1in, right=1.1in]{geometry}
\renewcommand{\baselinestretch}{1.1}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{mathtools}
\usepackage{gensymb}
\usepackage{url}

\newenvironment{smitemize}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}}
{\end{itemize}
}

\def\labelitemi{--}
\parindent=0pt

\begin{document}
\bibliographystyle{/Users/Lizzie/Documents/EndnoteRelated/Bibtex/styles/besjournals}
\renewcommand{\refname}{\CHead{}}

% This is doc, started for Mission Hill, but hopefully will eventually adapting to asking growers for data
\begin{center}
{\Large Data requested to build phenological models}
\end{center}
This document outlines data requested by the Wolkovich lab to build phenological models from budburst to sugar maturity. Please contact Lizzie (\url{e.wolkovich@ubc.ca}) questions or concerns.\\

{\bf Phenological data (including Brix)}\\

The most critical data are dates of the main phenological stages (budburst, flowering, veraison and sugar maturity). Ideally we hope for:
\begin{smitemize}
\item Budburst: green tips stage (if another stage close to this we can use, just be clear in what stage you are reporting)
\item 50\% flowering (we can use other percentages, just be clear in what stage you are reporting)
\item 50\% veraison (we can use other percentages, just be clear in what stage you are reporting)
\item Sugar maturity: The models benefit from \emph{all} Brix measurements (multiple time points of the same block/variety are very helpful). 
\end{smitemize}
For budburst, flowering and veraison, please indicate which percent of the buds and/or vines are at each stage (e.g., do you report 50\% flowering when most vines are at that stage? Or some other way?).\\

For each event we need to know the date, variety and vineyard block.\\

\emph{Varieties of special interest are:} Chardonnay, Merlot, Pinot Noir, Syrah, Riesling, Sauvignon blanc, Chasselas and Cabernet Sauvignon.\\

{\bf Climate data}\\
The models work best given the most local climate data, so any data for relevant vineyards you can share is helpful. We ideally want it from 1 September of the year {\bf before} budburst though the whole next season for each year with phenological data. We most critically need {\bf air temperature minima and maxima} but may try to use precipitation, humidity and other metrics if possible.\\

{\bf Yield data}\\
Vines carrying higher yields generally develop slower. Our models to try to account for this by modeling different vineyard blocks separately but actual yield data or metrics related to yield are most helpful for better models. If you can share any yield data we'd appreciate it.\\

{\bf Format of data}\\
The lab can re-format data, so the main thing we need is just the data in a way we can interpret. We do not want data formatting to be a major hurdle for sharing the data, so please let Lizzie know if it is and hopefully we can figure something out. We can read most file formats or pull data from AdCon etc. if you provide us access. Please feel free to ask any questions about this! 

\end{document}

\emph{References}
\bibliography{/Users/Lizzie/Documents/EndnoteRelated/Bibtex/LizzieMainMinimal}